Pilot study APEROL: questionnaire data

### 2021-01-21 studies/2020_Aperol_Git/questionnaire data/questionnaire_data_mandy.Rmd
update R script

# load tab delimited text file into R; 
# file should be in a subfolder named "data" for the script to select it
raw.data<-read.table("./data/short_questionnaire_data_1301211.txt",sep="\t", header=T)

# loading packages
# if a package is not installed on the current machine, it will install it
if (!require(tidyverse)) install.packages('tidyverse'); library(tidyverse)

@@

# Short Summary
This report provides a summary of the questionnaire data of the first pilot APEROL. The aim was to explore connections between the visual attention of the teacher and students’ in-class experience of the teachers’ "withitness". The study was conducted as a lab study with simulated scripted mini-lessons (10min per lesson). Data was collected from N = 8 teachers and N = 2 observers in 2 sessions. In each session, one subject took the role of teacher, while the other teachers acted as the class. After each lesson, the subjects answered items on classroom management, especially on the teacher's behaviour, using validated questionnaires. An external observer also answered a questionnaire. In addition, the teacher is asked to give a self-assessment on his/her classroom management.

### Data collection
For the pilot study, we collected questionnaire data from N = `r sum(n_distinct(raw.data$personID))` subjects in `r sum(n_distinct(raw.data$session))` sessions. The questionnaires were answered after each lesson (10min) by an observer, the teacher and the other teachers acting as the class.

@@

```{r overview demographics data table, echo=FALSE}
raw.data %>% # select data
  group_by(session, perspective) %>%
  summarise(N = n_distinct(personID),
            "mean age" = round(mean(age), 2),
            "min age" = min(age),
            "max age" = max(age)
            ) %>% 
  knitr::kable(digits = 2)

@@

### Plotting questionnaire data

```{r mean classroom management by session, echo = FALSE}

### classroom management session 1

# filter by parameter variable, i.e. create a subset for each session
km1.data <- raw.data %>% filter(scale == "Klassenmanagement",
                               session == "1")

view (km1.data)

# value sometimes contained text before filtering, we have to convert the numbers to numeric
km1.data <- km1.data %>%  mutate(value = as.numeric(as.character(factor(value))))

# create  mean, min, max, sd of classroom management
meankm1 <- mean(km1.data$value)
minkm1 <- min(km1.data$value)
maxkm1 <- max(km1.data$value)
sdkm1 <- sd(km1.data$value)
Nkm1 <- length(unique(km1.data$personID))
mdkm1 <- median(km1.data$value)

km1.data <- raw.data %>% filter(scale == "Klassenmanagement",
                               session == "1") %>% 
  group_by(perspective, target) %>%
  summarise(N = n_distinct(personID),
            "mean age" = round(mean(age), 2),
            "min age" = min(age),
            "max age" = max(age),
            "sd age" = round(sd(age), 2),
            ) %>% 
  knitr::kable(digits = 2)

### classroom management session 2

# filter by parameter variable, i.e. create a subset for each session
km2.data <- raw.data %>% filter(scale == "Klassenmanagement",
                               session == "2")

view (km2.data)

# value sometimes contained text before filtering, we have to convert the numbers to numeric
km2.data <- km2.data %>%  mutate(value = as.numeric(as.character(factor(value))))

# create  mean, min, max, sd of classroom management
meankm2 <- mean(km2.data$value)
minkm2 <- min(km2.data$value)
maxkm2 <- max(km2.data$value)
sdkm2 <- sd(km2.data$value)
Nkm2<-length(unique(km2.data$personID))

@@

The item "classroom management"  was rated by the observer, the teacher and students on average in session 1 with `r round(meankm1, 2)` (min = `r minkm1`; max = `r maxkm1`; SD = `r round(sdkm1,2)`; N = `r Nkm1`; lickert-scale 1-4; 4 = best evaluation) and in session 2 with `r round(meankm2, 2)` (min = `r minkm2`; max = `r maxkm2`; SD = `r round(sdkm2,2)`; N = `r Nkm2`; lickert-scale 1-4; 4 = best evaluation).

```{r Modul horicontal Plots,  fig.height = 4, message = FALSE, warning = FALSE , echo=FALSE}

# to create error bars, we need to summarise the data in a seperate dataframe
modul.plot.sd <-modul.plot.data %>%
  group_by(itemfulltext, Teilnehmende, .drop=TRUE) %>%
  summarise(mean = mean(value), n = n(), sd = sd(value), se = sd/sqrt(n), )

# and here comes the long plot
modul.plot<-ggplot(data = modul.plot.sd,
                           aes(x = itemfulltext, y = mean,
                               group=Teilnehmende, colour = Teilnehmende)) +
  geom_line()+
  geom_pointrange(position = position_dodge(0.1), aes(ymin = mean-se, ymax = mean+se))+
  theme_light() +
  ylim(1,5)+
  labs(data = modul.plot.data, y = "Ausprägung", x = NULL,
       title="Modulbewertung" ,
       subtitle=NULL)+
  theme(legend.position="bottom",
        panel.spacing.x = ,
        plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 55))+
  coord_flip()
modul.plot

@@

### Comparing Performance across age-groups and conditions
We used Wilcoxon/Mann-Whitney-U-tests to test for significant differences in children's performance across age-groups and conditions respectively.

xxx To Do xxx

### Full Analysis (GLMM, ANOVA)
We are going to analyze our outcome (correct choice 0/1) in a 2x2 design (condition (comm, noncomm); age (3yo, 4yo) using either a GLMM or ANOVA depending on prior tests on data structure and model stability.

xxx To Do xxx

@@

>>>>

# load tab delimited text file into R; 
# file should be in a subfolder named "data" for the script to select it
raw.data<-read.table("./data/short_questionnaire_data_2101211.txt",sep="\t", header=T)

# loading packages
# install.packages("stringr") # install stringr to use the filter() function with text data
library(stringr)

# if a package is not installed on the current machine, it will install it
if (!require(tidyverse)) install.packages('tidyverse'); library(tidyverse)

@@

# Short Summary
This report provides a summary of the questionnaire data of the first pilot APEROL. The aim was to explore connections between the visual attention of the teacher and students’ in-class experience of the teachers’ "withitness". The study was conducted as a lab study with simulated scripted mini-lessons (10min per lesson). Data was collected from N = 8 teachers in 2 sessions.

In each session, one subject took the role of teacher, while the other teachers acted as the class. After each lesson, the subjects answered items on classroom management, especially on the teacher's behaviour, using validated questionnaires. An external observer also answered a questionnaire. In addition, the teacher was asked to give a self-assessment on his/her classroom management.

### Data collection
For the pilot study, we collected questionnaire data from N = 8 subjects in `r sum(n_distinct(raw.data$session))` sessions. The questionnaires were answered after each lesson (10min) by an observer, the teacher and the three teachers acting as the class.

@@

```{r overview demographics data table, echo=FALSE}

# demographics, grouped by group (N, gender, meanage, max, min age)
raw.data %>%
  filter(what == "head") %>%
  group_by(group) %>%
  summarise(N=n(),
            "male" = sum(gender),
            "mean age" = round(mean(age), 2),
            "min age" = min(age),
            "max age" = max(age)
            ) %>% 
  knitr::kable(digits = 2)

@@

### Plotting scales

The questionnaire contains following scales:

(1) classroom management
(2) positive climate and motivation
(3) clarity and structuredness
(4) activation and support
(5) presence: posture and gaze
(6) presence: voice
(7) presence: verbal and non-verbal intervention
(8) turn taking
(9) natural behaviour

```{r cleaning items, echo = FALSE}

# to compare both sessions filter only Identical items
raw.data <- raw.data  %>% filter(str_detect(shortID, 
                                                 paste(c("km_mitbekommen",
                                                         "km_klar",
                                                         "km_ungestört",
                                                         "km_aktiv",
                                                         "lkm_freundlich",
                                                         "lkm_interesse",
                                                         "lkm_kritik",
                                                         "lkm_ausreden",
                                                         "lkm_überlegen",
                                                         "lkm_rückmeldungen",
                                                         "ks_funktion",
                                                         "ks_verständlich",
                                                         "ks_sichtbar",
                                                         "af_beiträge",
                                                         "af_wechselseitig",
                                                         "af_nachdenken",
                                                         "phb_stand",
                                                         "phb_augen",
                                                         "phb_blick",
                                                         "phb_vorsichgeht",
                                                         "phb_raum",
                                                         "phb_gestik",
                                                         "phb_alleangesehen",
                                                         "ps_deutlich",
                                                         "ps_klar",
                                                         "ps_impulse",
                                                         "pi_nonverbal",
                                                         "pi_zubewegen",
                                                         "pi_direkt",
                                                         "redeanteil",
                                                         "m_natürlich",
                                                         "m_fiktiv",
                                                         "m_verhalten"),
                                                    collapse = '|')))

@@

The scales on the quality of teaching are a validated questionnaire (Helmke 2014).Whereas the scales on the teacher's presence behaviour were derived from the research literature and were used in the pilot for the first time. 
For this reason, the various items for the scales ***presence*** are presented below.

### presence: posture and gaze

```{r mean presence: posture and gaze, echo = FALSE}

### presence: posture and gaze

# filter by parameter variable, i.e. create a subset for presence: posture and gaze
phb.data <- raw.data %>% filter(scale == "Präsenz/Haltung_Blick")

view (phb.data)

# value sometimes contained text before filtering, we have to convert the numbers to numeric
phb.data <- phb.data %>%  mutate(value = as.numeric(as.character(factor(value))))

# create  mean, min, max, sd of presence: posture/gaze
meanphb <- mean(phb.data$value)
minphb <- min(phb.data$value)
maxphb <- max(phb.data$value)
sdphb <- sd(phb.data$value)
Nphb <- phb.data %>% summarise(N = n_distinct(personID))

# filter by parametervariable, i.e. create a subset for experts and novice
expert.data <- phb.data %>% filter (group == "expert")
novice.data <- phb.data %>% filter (group == "novice")

# create mean, min, max, sd for expert/novice
# expert
expert.meanphb <- mean(expert.data$value)
expert.minphb <- min(expert.data$value)
expert.maxphb <- max(expert.data$value)
expert.sdphb <- sd(expert.data$value)
expert.Nphb<-length(unique(expert.data$personID))

#novice
novice.meanphb <- mean(novice.data$value)
novice.minphb <- min(novice.data$value)
novice.maxphb <- max(novice.data$value)
novice.sdphb <- sd(novice.data$value)
novice.Nphb<-length(unique(novice.data$personID))

@@

The mean for all items of the scale ***presence: posture and gaze*** is `r round(meanphb, 2)` (min = `r minphb`; max = `r maxphb`; SD = `r round(sdphb,2)`; N = `r Nphb`; lickert-scale 1-4; 4 = best evaluation).

The mini-lessons of the experts were evaluated on average with `r round(expert.meanphb, 2)` (min = `r expert.minphb`; max = `r expert.maxphb`; SD = `r round(expert.sdphb,2)`; N = `r expert.Nphb`).

The mini-lessons of the novices were evaluated on average with `r round(novice.meanphb, 2)` (min = `r novice.minphb`; max = `r novice.maxphb`; SD = `r round(novice.sdphb,2)`; N = `r novice.Nphb`).

```{r presence: posture and gaze line plots,  fig.height = 4, message = FALSE, warning = FALSE , echo=FALSE}

phb.data$group <- as.factor(phb.data$group)

# to create error bars, we need to summarize the data in a separate data frame
phb.plot.sd <- phb.data %>%
  filter (perspective == "teacher") %>%
  group_by(group, item.wordings, .drop=TRUE) %>%
  summarise(mean = mean(value), n = n(), sd = sd(value), se = sd/sqrt(n), )

# and here comes the long plot
phb.plot<-ggplot(data = phb.plot.sd,
                  aes(x = item.wordings, y = mean,
                      group = group, colour = group)) +
  geom_line()+
  geom_pointrange(position = position_dodge(0.1), aes(ymin = mean-se, ymax = mean+se))+
  theme_light() +
  ylim(1,4)+
  facet_grid()+
  labs(data = phb.data, y = "value", x = NULL,
       title="presence: posture and gaze" ,
       subtitle=NULL)+
  theme(legend.position="bottom",
        panel.spacing.x = ,
        plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 55))+
  coord_flip()
phb.plot

### Overview all scales

```{r mean classroom management, echo = FALSE}

### classroom management

# filter by parameter variable, i.e. create a subset for each session
km.data <- raw.data %>% filter(scale == "Klassenmanagement")

view (km.data)

# value sometimes contained text before filtering, we have to convert the numbers to numeric
km.data <- km.data %>%  mutate(value = as.numeric(as.character(factor(value))))

# create  mean, min, max, sd of classroom management
km.data.plot %>% km.data %>%

          meankm <- mean(km.data$value)
          minkm <- min(km.data$value)
          maxkm <- max(km.data$value)
          sdkm <- sd(km.data$value)
          Nkm <- km.data %>% summarise(N = n_distinct(personID))
          mdkm <- median(km.data$value)

### to do

- vergessen: Berufserfahrung bei Demographie mit aufnehmen
- selbst und fremdeinschätzung durch facet_grid bei plot hinzufügen

@@

Helmke, A., Helmke, T., Lenske, G., Pham, G., Praetorius, A. K., Schrader, F. W., & AdeThurow, M. (2014). Unterrichtsdiagnostik mit EMU. Aus-und Fortbildung der Lehrkräfte in Hinblick auf Verbesserung der Diagnosefähigkeit, Umgang mit Heterogenität und individuelle Förderung, 149-163.
